---
title: "Persistence times across studies"
author: "Jesse Brunner"
date: "10/18/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

# Studies with ATV at ASU in ~2002
```{r}
asu <- read_csv("Persistence_ATV_ASU2002.csv")
asu$Titer <- 10^asu$Log10Titer
qplot(Days, Titer, color=Trt, shape=Experiment, data=asu, geom="point") + 
  scale_y_log10() + 
  geom_smooth(method="lm") + 
  facet_grid(. ~ Experiment)
```
Not sure what to do with the first experiment... it seems super steep under either condition! I think I'll ignore this and just use the second experiment where we've got more data (note: from multiple ponds, combine). 

```{r}
qplot(Days, Titer, color=Trt, data=subset(asu, Experiment=="WaterDegII"), geom="point") + 
  scale_y_log10(breaks=10^(1:5)) + 
  geom_smooth(method="lm")

summary(lm(Log10Titer ~ Days, data=subset(asu, Experiment=="WaterDegII" & Trt=="Filtered")))
1/0.027413 # T90
0.027413/log10(exp(1)) # delta

summary(lm(Log10Titer ~ Days, data=subset(asu, Experiment=="WaterDegII" & Trt=="Unfiltered")))
1/0.16637 #T90
0.16637/log10(exp(1)) # delta
```



# Reinauer et al. 2005 data (from thesis via Rachel)
Reinauer, S., R. Bohm, and R. E. Marschang. 2005. Inactivation of tortoise viruses in the environment. Journal of Herpetological Medicine and Surgery 15:10-15.

Persistence in water
```{r}
rwater <- read_csv("Reinauer_Water.csv")

ggplot(rwater, aes(Day, LogTiter, color=factor(Temperature), shape=Replicate, linetype=Replicate)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  facet_grid(Substrate ~ GermCarrier)

# Tap water, 4C, rep A; First series
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="A" & Temperature==4)))
1/0.010691 #T90

# Tap water, 23C, rep A
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="A" & Temperature==23)))
1/0.05337 #T90

# Tap water, 37C, rep A
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="A" & Temperature==37)))
1/0.02373  #T90



# Tap water, 4C, rep B; Second Series
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="B" & Temperature==4)))
1/0.003969 #T90

# Tap water, 23C, rep B
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="B" & Temperature==23)))
1/0.11512 #T90

# Tap water, 37C, rep B
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="B" & Temperature==37)))
1/0.13864  #T90



# Tap water, 4C, no germ carrier, rep C; Second Series
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="C" & Temperature==4)))
1/0.0004711 #T90

# Tap water, 23C, no germ carrier, rep C
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="C" & Temperature==23)))
1/0.09007 #T90

# Tap water, 37C, no germ carrier, rep C
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="C" & Temperature==37)))
1/0.17783  #T90




# Pond water, 4C, rep D; First series
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="D" & Temperature==4)))
1/0.038321 #T90

# Pond water, 23C, rep D
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="D" & Temperature==23)))
1/0.032753 #T90

# Pond water, 37C, rep D
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="D" & Temperature==37)))
1/0.027213  #T90


# Pond water, 4C, rep E
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="E" & Temperature==4)))
1/0.01147 #T90

# Pond water, 23C, rep E
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="E" & Temperature==23)))
1/0.03858 #T90

# Pond water, 37C, rep E
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="E" & Temperature==37)))
1/0.04392  #T90


# Pond water, 4C, no germ carrier, rep F
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="F" & Temperature==4)))
1/0.06174 #T90

# Pond water, 23C, no germ carrier, rep F
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="F" & Temperature==23)))
1/0.21734 #T90

# Pond water, 37C, no germ carrier, rep F
summary(lm(LogTiter ~ Day, data=subset(rwater, Replicate=="F" & Temperature==37)))
1/0.20044  #T90
```
Filled in the values in the PersistenceTimes.csv spreadsheet

Now the persistence in dirt?
```{r}
soil <- read_csv("Reinauer_Dirt.csv")


ggplot(soil, aes(Day, LogTiter, color=Season)) + 
  geom_point() + 
  geom_smooth(method="lm", se=T) 

# In the soil; spring
summary(lm(LogTiter ~ Day, data=subset(soil, Season=="Spring")))
1/0.023539 #T90

# In the soil; summer
summary(lm(LogTiter ~ Day, data=subset(soil, Season=="Summer")))
1/0.3580 #T90
```



```{r}
persist <- read_csv("PersistenceTimes.csv")

#rate in log10 per day
persist$d <- 1/persist$T90
#rate in ln per day
persist$d <- persist$d/log10(exp(1))

# Refactor everything
persist$Study <- factor(persist$Study, levels = c("Brunner unpubl. data (ATV)", "Reinauer et al. 2005", "Nazir et al. 2012", "Johnson & Brunner 2014", "Munro et al. 2016"), ordered = TRUE)
persist$Activity <- factor(persist$Activity, levels = c("Microbial", "Enzymes (continual)", "Enzymes (initial)", "None"), ordered = TRUE)
persist$Matrix <- factor(persist$Matrix, levels = c("Substrate", "Water with substrate", "Water"), ordered=TRUE)
```

plot T90s against temperature by activity 
```{r}
ggplot(persist, aes(x=Temperature, y=T90, color=Activity, fill=Activity, shape=Study, linetype=Study)) + 
  geom_blank() + # required to keep legend in order
  
  geom_smooth(method="lm", formula = (y~I(1/x)), se=FALSE, alpha=0.25, size=0.5) +
  geom_smooth(data=filter(persist, Study != "Johnson & Brunner 2014"), 
              aes(shape=NULL, linetype=NULL), 
              method="lm", formula = (y~I(1/x)), se=FALSE) + 
  geom_point(size=2) + 
  facet_grid(. ~ Matrix, scales="free_x", space="free_x") + 
  scale_y_log10("T90 (days)", breaks=c(1,5,10,20,50,100,200,500,1000), minor_breaks=NULL) + 
  scale_x_continuous("Temperature (°C)", breaks=c(4,15,20,23,30,37), minor_breaks=NULL) +
    scale_shape_manual(values=LETTERS[1:5]) +
  scale_linetype_discrete() + theme_minimal() + 
  theme(legend.position="bottom", legend.direction = "vertical")

ggsave("T90.pdf", width=6.5, height=5)
```


change to rates of decline, $\delta$ in the model
```{r}

ggplot(persist, aes(x=Temperature, y=d, color=Activity, fill=Activity, shape=Study, linetype=Study)) + 
  geom_blank() + # required to keep legend in order
  geom_smooth(method = "lm", se=FALSE, size=0.5, alpha=0.25) +
  geom_smooth(data=filter(persist, Study != "Johnson & Brunner 2014"), 
              aes(shape=NULL, linetype=NULL), 
              method="lm", se=TRUE) + 
  geom_point(size=2.5) + 
  facet_grid(. ~ Matrix, scales="free_x", space="free_x") + 
  scale_y_sqrt(expression(paste("Rate of viral decay (", delta, "; per day)")),
               breaks=c(0.01, 0.05, 0.1, 0.2, 0.5, 1, 1.5, 2:4), minor_breaks=NULL) +
  scale_x_continuous("Temperature (°C)", breaks=c(4,15,20,23,30,37), minor_breaks=NULL) +
  coord_cartesian(ylim=c(0,4)) + 
  scale_shape_manual(values=LETTERS[1:5]) +
  scale_linetype_discrete() + theme_minimal() + 
  theme(legend.position="bottom", legend.direction = "vertical")

ggsave("Persistence.pdf", width=6.5, height=5)
```

